Breakpoint Models Show No Evidence of Recreational Adaptation to Increasing Wildfire Smoke in the American West

### 2021-07-07 MarkdownDocuments/Manuscript_Real.Rmd
fig captions drafted

# Introduction
Climate change has already begun to dramatically change the social-ecological landscapes in which humans have learned to operate [@IPCC]. All levels of human society must urgently adapt to these new realities if we are to continue to thrive over the coming millenia [@IPCC]. As a society, we are aware of this, but still are failing to respond adequately [@steg_limiting_2018]. Even when adaptive strategies and their benefits are known, we do not demonstrate adaptive behaviors [@van_valkengoed_meta-analyses_2019]. While research has shown that extreme events tend to spur climate adaptation at the government level, less is known about how individuals are reacting and possibly adapting to a changing climate [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].

A particularly salient environmental effect of climate change in the United States is the increasing wildfire frequency and severity over the last 40 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016]. As climate change accounts for approximately half of the increase in burned area during this time, fires are expected to worsen over the next 40 years as the climate continues to warm [@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015]. Accompanying the increase in wildfires in the American West, is a projected increase in wildfire smoke [@schroeder_validation_2008]. Fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].

![image alt text. Righ-hand y axis shows the standardized values of smoke ](../Figures/smkplt.png)

@@

![image alt text](../Figures/bpplot2.png)

@@

![image alt text](../Figures/ARVal.png)

@@

![image alt text](../Figures/slpz1.png)

![image alt text](../Figures/slpz2.png)

>>>>

# Introduction
Climate change has already begun to dramatically change the social-ecological landscapes in which humans have learned to operate [@IPCC]. All levels of human society must urgently adapt to these new realities if we are to continue to thrive over the coming decades[@IPCC]. As a society, we are aware of this, but still are failing to respond adequately [@steg_limiting_2018]. Even when adaptive strategies and their benefits are known, we do not demonstrate adaptive behaviors [@van_valkengoed_meta-analyses_2019]. While research has shown that extreme events tend to spur climate adaptation at the government level, less is known about how individuals are reacting and possibly adapting to a changing climate [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].

A particularly salient environmental effect of climate change in the United States is the increasing wildfire frequency and severity over the last 40 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016]. As climate change accounts for approximately half of the increase in burned area during this time, fires are expected to worsen over the next 40 years as the climate continues to warm [@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015]. Accompanying the increase in wildfires in the American West, is a projected increase in wildfire smoke [@schroeder_validation_2008]. Fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014]. referred to

![Smoke observations in each of the 32 national parks included in this study from 1980-2019. Smoke events greater than 5 on our standardized scale (10 standard deviations above the mean) are labeled. The right-hand y-axis shows the standardized values of smoke referred to when describing the chosen critical thresholds. Smoke values on the left-hand y-axis are....ALEX..](../Figures/smkplt.png)

@@

![Conceptual diagram showing the design of the three breakpoint models used in this analysis. Note that slope 1 is not necessarily 0 in the design of these models. All three models also do not necessarily have different intercepts. The lines showing the three different breakpoints are separated along the y-axis only for visual clarity.](../Figures/bpplot2.png)

@@

![Scatterplot showing the within-sample predictive accuracy of the autoregressive baseline model used in this study. This model accounts for 40% of all variation in the data.](../Figures/ARVal.png)

@@

![Posterior parameter estimates for slope 1 for each of the 32 national parks included in this study for each of the three breakpoint models. Points represent the median prediction. Thick lines show the 50% credibility intervals and thin lines show the 90% intervals.](../Figures/slpz1.png)

![Posterior parameter estimates for slope 2 for each of the 32 national parks included in this study for each of the three breakpoint models. Points represent the median prediction. Thick lines show the 50% credibility intervals and thin lines show the 90% intervals.](../Figures/slpz2.png)
